Recent demography drives changes in linked selection across the maize genome

### 2015-11-10 manuscript.tex
stupid pnastwo

The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
Comparisons of estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences.  
One possibility for example, is that if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role. 
Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.  
Our results, for example, are not very different from those of \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions and estimate that as many as 80\% of adaptive substitutions occur outside of genes.  
Future comparative analyses using a common statistical framework (e.g. \cite{corbett2015}) should allow explicit testing of this idea.

\subsection{Demography influences the efficiency of purifying selection} %JRI NOT DONE
One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a demestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.

@@

%We propose that this observation could also result from more efficient selection against such mutations. This proposition is in agreement not only with our maize observations, but also with the results of \cite{}, who showed via simulations that the increased efficiency of selection in a larger population can partially offset the increase in the total number of segregating deleterious variants.
%While the large numbers of rare variants present in expanding populations may contribute significantly to the genetic basis of complex disease \cite{keinan2012}, recent analyses suggest that 
%. However, more efficient purifying selection in larger populations, which we show operates specifically on these rare variants, should decrease genetic load by purging rare deleterious variants.

%Across species, our results demonstrate that the genetic impacts of major demographic events such as bottlenecks depend not only on population size reductions, but also on subsequent, potentially rapid and recent, expansion that often comes afterwards. Reduced population sizes do lead to a decreased efficiency of selection \cite[e.g]{renaut2015}, which has long-term impacts on diversity measured via $\pi$. But new mutations, which we measure with singleton diversity, are oblivious to past demography, and therefore selection for or against such sites depends primarily on recent demographic processes.}

\section{Demography and selection interact to shape genetic diversity} %JRI NOT DONE
Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection.

\cite{Elyashiv:2010ga} $N_e$ among populations leads to differences in efficiency of purifying selection

>>>>

The absence of evidence for a genome-wide impact of hard sweeps in coding regions differs markedly from observations in \emph{Drosophila} \cite{sattath2011} and \emph{Capsella} \cite{williamson2014}, but is consistent with data from humans \cite{hernandez2011,pritchard2010genetics}.
Comparisons of estimates of the percentage of nonsynonmyous substitutions fixed by natural selection \cite{ross2009,sella2009,eyre2009estimating,williamson2014} give similar results. 
While larger long-term $N_e$ likely explains some of this difference, we see little change in the importance of hard sweeps in genes in singleton diversity in modern maize (Figure \ref{sFig:singleton}, perhaps suggesting other factors may contribute to these differences.  
One possibility for example, is that if mutational target size scales with genome size, the larger genomes of human and maize may offer more opportunities for noncoding loci to contribute to adaptation, with hard sweeps on nonsynonymous variants then playing a relatively smaller role. 
Support from this idea comes from numerous cases of adaptive transposable element insertion modifying gene regulation in maize \cite{studer2011,castelletti2014mite,mao2015,yang2013} and studies of local adaptation that show enrichment for SNPs in regulatory regions in teosinte \cite{pyhajarvi2013} and humans \cite{fraser2013gene} but for nonsynonymous variants in the small \emph{Arabidopsis} genome.  
Our results, for example, are not very different from those of \cite{Halligan:2013ir}, who find no differences in diversity around nonsynonymous and synonymous substitutions and estimate that as many as 80\% of adaptive substitutions occur outside of genes.  
Future comparative analyses using a common statistical framework (e.g. \cite{corbett2015}) and considering additional ecological and life history factors (c.f. \cite{leffler2012}) should allow explicit testing of this idea.

\subsection{Demography influences the efficiency of purifying selection} %JRI NOT DONE
One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
The efficiency of purifying selection is proportional to the effective population size \cite{kimura1984}, and these results are thus consistent with a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} and our demographic analyses showing rapid expansion and a much larger modern $N_e$.

@@

\section{Rapid changes in linked selection} %JRI NOT DONE
Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
While a number of authors have incorporated demographic change into methods for detecting selection \cite{eyre2009estimating} \jri{include other sites. maybe zeng and charlesworth 2010 genetics?, XP-CLR? others?} and others have compared estimates of the efficiency of selection across species \citex or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size

 long-term estimates of the effective population size or

 cite e.g. 
 $N_e$ among populations leads to differences in efficiency of purifying selection
